---
title: "How to make a particularly cool plot with ggplot"
author: "Xinrui Chen"
date: "October 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, tidy=FALSE,error=FALSE, message=FALSE, warning=FALSE, tidy.opts=list(width.cutoff=60))
```

## Introduction

Exploratory data analysis is a loosely defined process. It is roughly the part between loading data in and the formal analysis you will perform. It includes (but is not limited to)

  * Visualization
  * Checks for data completeness and reliability
  * Quantification of variability
  * Initial evaluation of hypothesis / hypothesis generation

R has several systems for making graphs, but ggplot2 is one of the most elegant and most versatile. ggplot2 implements the grammar of graphics, a coherent system for describing and building graphs.



## Good plots overview
Think about the story your plot is telling. What will someone learn while looking at your plot? Some practical things to make a plot better.

  * Label the axis
  * Have a title
  * Make sure the title and axis are large enough
  * Label the plot or have a legend
  * Text on plots is great!
  * Use color and fill effectively to differentiate groups
  * Don’t make a plot overly complex – if it is getting too complex perhaps make two plots
  * Get rid of non-essentials

## ggplot2

In ggplot2,

  * the ggplot() command creates a plot, with the argument data telling us which data to use
  * the aes() command within ggplot() indicates the desired x and y variables
  * plots are built with elements
  * the element geom_point() indicates that we are using a scatterplot

## Prerequisites
To access the datasets, help pages, and functions that we will use in this chapter, load the packages by running this code:

```{r}
library(tidyverse)
library(ggplot2)
library(mdsr)
```

## A Graphing Template
To make a graph, replace the bracketed sections in the following code with a dataset, a geom function, or a collection of mappings:

```{r，eval = FALSE}
ggplot(data = <DATA>) +
<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```


## Example to explore ggplot2

We will be using the data set CIACountries from the mdsr package. The CIACountries data table contains the following seven variables for 236 countries:

  * population (pop)
  * area (area)
  * gross domestic product (gdp)
  * percentage of GDP spent on education (educ)
  * length of roadways per unit area (roadways)
  * internet use as a fraction of the population (net_users)
  * number of barrels of oil produced per day (oil_prod)

```{r}
head(CIACountries)
```

### Basic plots

```{r}
g <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) +
theme(axis.text=element_text(size=10),
axis.title=element_text(size=10),
plot.title=element_text(size = 14)) +
ylab('Gross Domestic Product') +
xlab('Percentage of Gross Domestic \n Product Spent on Education') +
ggtitle('Gross Domestic Product and Education')
```

Scatterplot using only the position aesthetic for glyphs:

```{r}
g+ geom_point()
```

Scatterplot in which net_users is mapped to color:

```{r}
g + geom_point(aes(color = net_users), size = 3) +
guides(color=guide_legend(title="Internet Use"))
```

Scatterplot using both location and label as aesthetics:

```{r}
g + geom_text(aes(label = country, color = net_users), size = 3) +
guides(color=guide_legend(title="Internet Use"))
```

Scatterplot in which net_users is mapped to color and educ mapped to size:

```{r}
g + geom_point(aes(color = net_users, size = roadways)) +
guides(color=guide_legend(title="Internet Use"),
size=guide_legend(title='Roadway'))
```

### Scale
From the graphs that we made it is difficult to discern differences in GDP because there is a right-skew in GDP and we made a choice to use a linear scale on our plots.

  * A logarithmic scale would make the plot much more readable!
  * We will use the coord_trans() function to put our data on a logarithmic scale.


Scatterplot using the logarithmic transform of GDP that helps to mitigate visual clustering caused by the right-skewed distribution of GDP among cities:

```{r}
g + geom_point(aes(color = net_users, size = roadways)) +
guides(color=guide_legend(title="Internet Use"),
size=guide_legend(title='Roadway')) +
coord_trans(y = "log10") # rescale the y-axis value
```

As an alternative to coord_trans() we can use the scale_y_continuous() function. The only difference is where the tick marks on the axis are drawn. The functions scale_x_continuous(), scale_x_discrete(), and scale_color() all perform analogous operations on different aesthetics.

Scatterplot using the logarithmic transform of GDP that helps to mitigate visual clustering caused by the right-skewed distribution of GDP among cities:

```{r}
g + geom_point(aes(color = net_users, size = roadways)) +
guides(color=guide_legend(title="Internet Use"),
size=guide_legend(title='Roadway')) +
scale_y_continuous(name = 'Gross Domestic Product', trans = 'log10')
```

### Facets

Using multiple aesthetics such as shape, color, and size to display multiple variables can produce a really hard to read graph! Facets are multiple side by side graphs used to display levels of categorical variables which can help.

Scatterplot using facets for different ranges of Internet connectivity:

```{r}
g + geom_point(alpha = 0.9, aes(size = roadways)) +
coord_trans(y = 'log10') +
facet_wrap(~net_users, nrow = 1) +
theme(legend.position = 'top') +
theme(axis.text.x =element_text(size=10))
```

### Summary: the Layered Grammar of Graphics

```{r，eval = FALSE}
ggplot(data = <DATA>) +
<GEOM_FUNCTION>(
mapping = aes(<MAPPINGS>),
stat = <STAT>,
position = <POSITION>
) +
<COORDINATE_FUNCTION> +
<FACET_FUNCTION>
```

Our new template takes seven parameters, the bracketed words that appear in the template. In practice, you rarely need to supply all seven parameters to make a graph because ggplot2 will provide useful defaults for everything except the data, the mappings, and the geom function. The seven parameters in the template compose the grammar of graphics, a formal system for building plots. The grammar of graphics is based on the insight that you can uniquely describe any plot as a combination of a dataset, a geom, a set of mappings, a stat, a position adjustment, a coordinate system, and a faceting scheme.


