---
title: "How to make plots with plotly"
author: "Xinrui Chen"
date: "October 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, tidy=FALSE,error=FALSE, message=FALSE, warning=FALSE, tidy.opts=list(width.cutoff=60))
```

## Introduction

Plotly is a technical computing company headquartered in Montreal, Quebec, that develops online data analytics and visualization tools. Plotly provides online graphing, analytics, and statistics tools for individuals and collaboration, as well as scientific graphing libraries for Python, R, MATLAB, Perl, Julia, Arduino, and REST.

In this post, we will discuss plotly for R with examples.

## Prerequisites
Load the plotly package.

```{r}
library(plotly)
```

## Scatter Plot

We will use iris data set to explore scatter plot.

```{r}
head(iris)
```


```{r}
p <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length)
p
```

### Styled Scatter Plot
Then we change the layout of the basic scatter plot.

```{r}
p <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length,
        marker = list(size = 10,
                       color = 'rgba(255, 182, 193, .9)',
                       line = list(color = 'rgba(152, 0, 0, .8)',
                                   width = 2))) %>%
  layout(title = 'Styled Scatter',
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))

p
```



### Qualitative Colorscales
```{r}
p <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species)
p
```

### ColorBrewer Palette Names
```{r}
p <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, colors = "Set1")
p
```

### Custom Color Scales
```{r}
pal <- c("red", "blue", "green")

p <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, colors = pal)
p
```

### Data Labels on Hover
```{r}
d <- diamonds[sample(nrow(diamonds), 1000), ]

p <- plot_ly(
  d, x = ~carat, y = ~price,
  # Hover text:
  text = ~paste("Price: ", price, '$<br>Cut:', cut),
  color = ~carat, size = ~carat
)
p
```


## Plotting Markers and Lines

```{r}
trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

p <- plot_ly(data, x = ~x) %>%
  add_trace(y = ~trace_0, name = 'trace 0',mode = 'lines') %>%
  add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers') %>%
  add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')
p
```

## Bar charts
```{r}
p <- plot_ly(
  x = c("giraffes", "orangutans", "monkeys"),
  y = c(20, 14, 23),
  name = "SF Zoo",
  type = "bar"
)
p
```

## Pie charts
```{r}
USPersonalExpenditure <- data.frame("Categorie" = rownames(USPersonalExpenditure), USPersonalExpenditure)
data <- USPersonalExpenditure[, c('Categorie', 'X1960')]

colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)')

p <- plot_ly(data, labels = ~Categorie, values = ~X1960, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        text = ~paste('$', X1960, ' billions'),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
                      #The 'pull' attribute can also be used to create space between the sectors
        showlegend = FALSE) %>%
  layout(title = 'United States Personal Expenditures by Categories in 1960',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
p
```

## Boxplots
```{r}
p <- plot_ly(ggplot2::diamonds, y = ~price, color = ~cut, type = "box")
p
```

## Polar charts
```{r}
p <- plot_ly(
  type = 'scatterpolar',
  r = c(0,1,2,2),
  theta = c(0,45,90,0),
  mode = 'markers'
)
p
```



In summary, plotly is very useful in interactive visualization.Apart from above plots, we can also draw 3D plots, map plots and many other plots with plotly.Try it by yourself!